---
output: github_document
---
# TCT 

The goal the TCT package is to provide a set of easy to use functions to
perform time components test. ## Installation You can install the development
version of TCT from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("florianstijven/TCT")
```
## Example 

This is a basic example which shows you how to use the time-component
test methodology in combination with an estimated MMRM model. First the MMRM is
estimated. All we need from this estimated model is the estimated mean vector
and estimated variance-covariance matrix for this estimated vector.

```{r example, message=FALSE}
library(TCT)
library(dplyr)
# Example data set transformed to format required by TCT()
data_test = simulated_test_trial %>%
  mutate(
    time_int = (Week %/% 25) + 1,
    arm_time = ifelse(time_int == 1L,
                      "baseline",
                      paste0(arm, ":", time_int))
  )
# Fit a MMRM model to the data set. The parameter estimates of this model form
# the basis to perform the time component tests.
mmrm_fit = analyze_mmrm(data_test)
```

After fitting a MMRM model, we can use the `TCT()` function to perform the
time-component test. The first step is to estimate the slowing factor for each
time point.

```{r cars}
# TCT() is the main function for performing time component tests. The estimated
# mean parameters of the MMRM model and associated variance-covariance matrix
# form the basis for the time component test.
TCT_fit = TCT(time_points = 0:4,
              ctrl_estimates = coef(mmrm_fit)[c(9, 1:4)],
              exp_estimates = coef(mmrm_fit)[5:8],
              vcov = vcov(mmrm_fit)[c(9, 1:4, 5:8), c(9, 1:4, 5:8)],
              interpolation = "spline",
              B = 1e4)
# Summary of the results of the time component test. Here we have estimated a
# treatment effect summary(x = TCT_fit) parameter on the time scale for every
# measurement occasion. 
summary(TCT_fit)

```

The next step is to estimate a common slowing factor. 

```{r}
# The object returned by TCT() can be used as argument to the TCT_common() function.
# This function estimates a single common slowing factor.
TCT_common_fit = TCT_common(TCT_fit, B = 1e4, bs_fix_vcov = FALSE)
summary(TCT_common_fit)
```
