---
output: github_document
---
# TCT 

The goal the TCT package is to provide a set of easy to use functions to
perform time components test. 

## Installation You can install the development

version of TCT from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("florianstijven/TCT")
```
## Example 

This is a basic example which shows you how to use the Meta-TCT methodology in
combination with an estimated MMRM model. First the MMRM is estimated. All we
need from this estimated model is the estimated mean vector and estimated
variance-covariance matrix for this estimated vector.

```{r example, message=FALSE}
library(TCT)
library(dplyr)
# Example data set transformed to format required by TCT()
data_test = simulated_test_trial %>%
  mutate(
    time_int = (Week %/% 25) + 1,
    arm_time = ifelse(time_int == 1L,
                      "baseline",
                      paste0(arm, ":", time_int))
  )
# Fit a MMRM model to the data set. The parameter estimates of this model form
# the basis to perform the time component tests.
mmrm_fit = analyze_mmrm(data_test)
```

After fitting a MMRM model, we can use the `TCT_meta()` function to perform the
time-component test. The first step is to estimate the slowing factor for each
time point. Different approaches for inference are available. Note that the
estimators in these different approaches are equivalent, but the corresponding
measures of uncertainty are not equivalent. Information on these approaches can
be found in the function documentation.

```{r cars}
# TCT_meta() is the main function for performing time component tests. The estimated
# mean parameters of the MMRM model and associated variance-covariance matrix
# form the basis for the time component test.
TCT_fit = TCT_meta(
  time_points = 0:4,
  ctrl_estimates = coef(mmrm_fit)[c(9, 1:4)],
  exp_estimates = coef(mmrm_fit)[5:8],
  vcov = vcov(mmrm_fit)[c(9, 1:4, 5:8), c(9, 1:4, 5:8)],
  interpolation = "spline",
  inference = "score",
  B = 1e4
)
# Summary of the results of the time component test. Here we have estimated a
# treatment effect parameter on the time scale for every measurement occasion.
summary(TCT_fit)
```

The next step is to estimate a common slowing factor. As before, different
approaches to inference are available. However, the estimators in these
different approaches are no longer equivalent. In the next code chunk,
score-based weighted with adaptively selected weights is illustrated. These
weight the different measurement occasions such the the variance of the
corresponding estimator is minimized.

```{r}
# The object returned by TCT() can be used as argument to the TCT_common() function.
# This function estimates a single common slowing factor.
TCT_common_fit = TCT_meta_common(TCT_fit,
                                 B = 0,
                                 inference = "score",
                                 type = "custom")
summary(TCT_common_fit)
```
