---
title: "TCT_meta"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TCT_meta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of the this vignette is to show the `TCT_meta()` and `TCT_meta_common()`
functions can be used for conducting Time-Component Test using only summary 
information. 

```{r setup}
# The TCT R-package is loaded.
library(TCT)
library(dplyr)
```

# Data

The usage of the previously mentioned functions is illustrated with the 
`simulated_test_trial` data available after loading the `TCT` package. These 
data have been simulated to represent a typical trial in Alzheimer's disease. 
A description of this data set can be accessed through `?simulated_test_trial`.
For completeness, the most important features of these data are summarized in this
vignette:

* There are 5 measurement occasions, one of which is at baseline. There are thus 4
post-randomization time points at which the acceleration factor can be
estimated.
* The total sample size is 1000 with 500 patients in each treatment arm. This
corresponds to a rather large trial.
* The patients are assumed to be followed for 100 weeks following randomization.
In real-life data sets, there would obviously be missing data and possibly
truncation by death. These potential issues are ignored, but should be dealt
with appropriately in any real-life application of the TCT methodology.
* The "true" acceleration factor is `1/1.3` in these data. Acknowledging the
sampling variability, the estimated acceleration factor should be close to this
value.


# MMRM

As previously mentioned, the meta-TCT methodology only requires summary-level
information. In principle, this information can come from a wide variety of
longitudinal models such as linear mixed models, generalized linear mixed
models, and quantile regression models. In this vignette, the methodology will 
be illustrated using the linear mixed model. Specifically, we will use the 
so-called mixed-model for repeated measures (MMRM) as implemented in the `mmrm`
R-package. 

The summary level information should contain two components:

1. Estimated mean outcome at each measurement occasion. It is further assumed
that there is only 1 common estimated for the mean at baseline. Note that the 
meta-TCT methods rely on asymptotic normality of these estimators.
2. Estimated variance-covariance matrix for the estimated vector of mean
parameters.

Note that the mean can be replaced with any other function of the data such as 
quantiles.

In the following code snippet, the MMRM is fitted to the data. Note that the
model should be parameterized in such a way that estimates of the mean at each
measurement occasion are obtained. This is not the conventional parameterization
where the control means and mean differences are the parameters being estimated.

```{r}
# The data are fitted using the a built-in function that essentially is a
# wrapper for the mmrm::mmrm() function.

# The data are first transformed into the required format. Specifically, we
# construct an arm_time variable which represents each arm-measurement occasion
# combination, except for the baseline measurement. For the latter, only a
# single category, common to both treatment arms, is present. This ensures that
# there is only a single mean parameter for that occasion. 
data_test = simulated_test_trial %>%
  dplyr::mutate(
    time_int = (Week %/% 25) + 1,
    arm_time = ifelse(time_int == 1L,
                      "baseline",
                      paste0(arm, ":", time_int))
  )
# Fit a MMRM model to the data set. The parameter estimates of this model form
# the basis to perform the time component tests.
mmrm_fit = analyze_mmrm(data_test,
                        package = "mmrm",
                        method = "REML",
                        type = "full")
```

We can print the summary of this fitted model. However, the mean estimates may
not be informative with respect to how large the treatment effect actually is. 
In contrast, an estimate of the treatment effect on the vertical scale is 
much more informative. This is the goal of the TCT methodology.

```{r}
summary(mmrm_fit)
```



# Meta-Analytic Time-Component Test

## Exctraction of Summary Information

The first step is to extract the relevant summary information from the fitted
models. Here, we show how this can be done using the fitted MMRM model. This
step can easily be modified for other types of models, or reported estimates can
be used (which does not require the data analyst to have access to
individual-patient data).

We first print the estimated mean parameters. The names of the elements of this
vector indicate to which measurement occasion these estimates correspond. This
information is needed to know which estimates should be used to construct the
reference trajectory, and which estimates should be mapped to the reference
trajectory.

```{r}
print(coef(mmrm_fit))
```

We now see that the 9th estimate corresponds to baseline, and the 1th to 4th 
estimates correspond to the subsequent estimates in the control group. These 
estimates are added to the vector, `control_estimates`, which will be used to
construct the reference trajectory.

```{r}
control_estimates = coef(mmrm_fit)[c(9, 1:4)]
```

The 5th to 8th estimates correspond to the post-randomization measurement
occasions in the active treatment group. These estimates are added to the vector,
`active_estimates`, the elements of which will be mapped to the reference 
trajectory. 

```{r}
active_estimates = coef(mmrm_fit)[c(5:8)]
```

Finally, we need to extract the estimated variance-covariance matrix. Extra care
should be taken here to ensure that the rows and columsn correspond to the
estimates in the following order: `c(control_estimates, active_estimates)`. To
be sure that everything is correct, we also print the column and row names of
the extracted matrix.

```{r}
vcov_mmrm = vcov(mmrm_fit)[c(9, 1:4, 5:8), c(9, 1:4, 5:8)]
colnames(vcov_mmrm)
rownames(vcov_mmrm)
```


## Time-Specific Acceleration Factors

The second step is to estimate the time-specific acceleration factors. These
estimates are based on the mapping of the estimates in `active_estimates` onto
the trajectory based on the estimates in `control_estimates`. In this second
step, a different acceleration factor is estimated for each measurement
occasion.

The `TCT_meta()` function performs this step. This function requires two sets of
arguments:

1. The first set of arguments pertains to the actual summary information. The
`ctrl_estimates`, `exp_estimates`, and `vcov` arguments require the estimated
means and corresponding variance-covariance matrix. In addition, we need to
specify the timing of the estimates in `control_estimates` through the
`time_points` argument. In the simulated data, these are 5 equally spaced
measurement occasions, so we can give the `0:4` vector. Multiplying this vector
with a constant does not change the results. Clearly, if the occasions are not
equally-spaced, this should be reflected in `time_points`.
2. The second set of arguments pertains to options for inference. All options 
are described in the function documentation. 
  * For constructing the reference trajectory, we use cubic spline interpolation,       hence, the `"spline"` argument for `interpolation`.
  * For inference on the estimated acceleration factors, we use the `"score"` based
    inference.
  * `B` specifies the number of bootstrap replications. We do not use the bootstrap     for inference in this vignette. 
  
```{r}
TCT_meta_fit = TCT_meta(
  time_points = 0:4,
  ctrl_estimates = control_estimates,
  exp_estimates = active_estimates, 
  vcov = vcov_mmrm, 
  interpolation = "spline", 
  inference = "score", 
  B = 0
)
```

We can call the `summary()` method on the object returned by `TCT_meta()` to 
get more information on the estimates and corresponding inference.

```{r}
summary(TCT_meta_fit)
```

This shows that there is some evidence of slowing at each measurement occasion.
In addition, the acceleration factor seems quite constant across these
measurement occasions. Therefore, we proceed by estimating a common acceleration
factor.

## Common Acceleration Factor

The third (and optional) step in the TCT analysis is estimating the common
acceleration factor. This estimate will only make sense of there is no strong
evidence against a common acceleration factor. Above, we have not seen any
evidence against a common acceleration factor, so we can safely continue.

This step is implemented in the `TCT_meta_common()` function. This function
takes the object returned by `TCT_meta()` as first argument. The remaining
arguments specify options for inference. Note that of the inference options used
in `TCT_meta()`, only `interpolation` has any effect on the behavior of
`TCT_meta_common()`. Only the inference options specified in `TCT_meta_common()`
are important.

Whereas there are only a limited number of inference options in the case of 
time-specific acceleration factors, there are many more options for the case
of a common acceleration factor. 

* The `inference` argument specifies the method of inference. We choose once `"score"` and
once `"least-squares"`.
  * For `"score"` based inference, several `type`s are possible. We choose 
   `type = "custom"`. If we then leave the `weights` unspecified, the weights
   will be adaptively selected as to minimize the standard error of the common 
   acceleration factor estimate.
* As before, `B` specifies the number of replications in a bootstrap. We do not use
the boostrap, so we set `B = 0`.
* `select_coef` is an optional argument which allows use to only use a subset of
the estimates in `control_estimates` for estimating a common acceleration
factor. This is useful when the is, for example, evidence of non-proportional
slowing at the first measurement occasions. By default, all measurement
occasions are used.


```{r}
TCT_common_score = TCT_meta_common(
  TCT_Fit = TCT_meta_fit, 
  inference = "score", 
  B = 0,
  type = "custom"
)
TCT_common_ls = TCT_meta_common(
  TCT_Fit = TCT_meta_fit, 
  inference = "least-squares", 
  B = 0
)
```

We can again use the `summary()` method on the object returned by `TCT_meta()` to 
get more information on the estimates and corresponding inference.

```{r}
summary(TCT_common_score)
summary(TCT_common_ls)
```

Both estimates are fairly close to each other, and to the expected value.
However, the p-values lead to different conclusions regarding significance. This
is not unexpected as both methods approach the problem from a different
perspective.

In any case, we can conclude that there is borderline evidence of a treatment 
effect. In addition, and this is the main contributing of the TCT methodology, 
we can conclude that the treatment slows the disease process with maximally 
30-35%. 







